---
title: "Lab 5A"
output: html_document
theme: readable
---

```{r, message=FALSE}
library(tidyverse)

```



# Boxplot

```{r}
surveys_complete <- read_csv("Results/surveys_complete.csv")

ggplot(data=surveys_complete, mapping=aes(x=species_id, y= weight))+
  geom_boxplot()
```


```{r}
ggplot(data=surveys_complete, mapping=aes(x=species_id, y= weight))+
  geom_boxplot(alpha=0)+
  geom_jitter(alpha=0.3, color= "tomato")
```

What do you need to change inthe code to put the boxplot in front of the points such that it’s not hidden?



```{r}
ggplot(data=surveys_complete, mapping=aes(x=species_id, y= weight))+
  geom_jitter(alpha=0.3, color= "tomato")+
  geom_boxplot(alpha=0)
```

## **Challenges**

Replace the box plot with a violin plot; see geom_violin()

```{r}
ggplot(data=surveys_complete, mapping=aes(x=species_id, y= weight))+
  geom_jitter(alpha=0.3, color= "tomato")+
  geom_violin()
```



Represent weight on the logscale; see scale_y_log10()


```{r}
ggplot(data=surveys_complete, mapping=aes(x=species_id, y= weight))+
  geom_jitter(alpha=0.3, color= "tomato")+
  geom_violin()+
  scale_y_log10()
```

Create boxplot for hindfoot_length. Overlay the boxplot layer on a jitter layer to show actual measurements

```{r}
ggplot(data=surveys_complete, mapping = aes(x= species_id, y=hindfoot_length))+
  geom_jitter( aes(color=as.factor(plot_id)), alpha=0.3)+
  geom_boxplot(alpha=0)
```
Hint: Check the class for
plot_id
. Consider changing the class of
plot_id
from integer to factor. Why does this change how R makes the graph?

```{r, echo=FALSE}
print("When plot_id is in the class numeric R assign colors to plot_id as a gradient. But When it is a factor each factor is colored as a distinct category")
```

# Ploting time series data

Calculate number of counts per year for each genus

```{r}
yearly_counts <- surveys_complete %>% 
 count(year,genus)
```

```{r}
ggplot(data=yearly_counts, aes(x=year, y=n))+
  geom_line()
```
```{r}
ggplot(data=yearly_counts, aes(x=year, y=n))+
  geom_line(aes(group=genus))
```
```{r}
ggplot(data=yearly_counts, aes(x=year, y=n))+
  geom_line(aes(group=genus, color=genus))
```

# Integrating the pipe operator with ggplot2

```{r}
yearly_counts %>% 
ggplot(aes(x=year, y=n))+
  geom_line(aes(group=genus, color=genus))
```
```{r}
yearly_counts_graph <- surveys_complete %>% 
 count(year,genus) %>%
 ggplot(aes(x=year, y=n))+
  geom_line(aes(group=genus, color=genus))

yearly_counts_graph
```

# Faceting

```{r}
ggplot(data=yearly_counts, aes(x=year, y=n))+
  geom_line(aes(group=genus, color=genus))+
  facet_wrap(facets = vars(genus))
```

```{r}
yearly_sex_counts <- surveys_complete %>% 
  count(year, genus, sex)
```

```{r}
ggplot(data=yearly_sex_counts, mapping = aes(x=year, y= n))+
  geom_line(aes(color= sex))+
  facet_wrap(facets = vars(genus))
```
```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(sex), cols =  vars(genus))
```

```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(genus))
```

```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(cols =  vars(genus))
```


# ggplot2 themes

```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(cols =  vars(genus))+
  theme_bw()
```

## **Challenge**


Use what you just learned to create a plot that depicts how the average weight of each species changes through the years.

```{r, fig.width=10}
surveys_complete %>% 
  group_by(species, year) %>% 
  summarise(Mean_weight= mean(weight)) %>% 
  ggplot(aes(x=year, y=Mean_weight))+
  geom_line(aes(color=species))+
  facet_wrap(facets = vars(species), scales = "free_y")+
  theme_classic()
  
```

```{r, fig.width=10}
ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus))+
  labs(title = "Observed genera through time",
       x="Year of observation",
       y= "Number of individuals")+
  theme_classic()+
  theme(text = element_text(size=16))
```

```{r, fig.width=10}
ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus))+
  labs(title = "Observed genera through time",
       x="Year of observation",
       y= "Number of individuals")+
  theme_classic()+
  theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
        axis.title.y = element_text(colour = "grey20", size = 12),
        strip.text = element_text(face="italic"),
        text = element_text(size=16))
```
```{r}
grey_theme <- theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
        axis.title.y = element_text(colour = "grey20", size = 12),
        strip.text = element_text(face="italic"),
        text = element_text(size=16))

ggplot(surveys_complete, aes(x=species_id, y= hindfoot_length))+
  geom_boxplot()+
  grey_theme
```

## **Challenge**
Improve one of the plots generated in this exercise or create a beautiful graph of your own.

```{r, fig.width=10}
ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus), scales = "free_y")+
  labs(title = "Observed genera through time",
       x="Year of observation",
       y= "Number of individuals")+
#scale_color_discrete(name="Sex of Genus", labels = c("F"="Female","M"="Male"), aesthetics=c("color"))+
  scale_color_manual( breaks =c("F","M"), labels=c("Female","Male") ,values=c("#000000", "#E69F00"))+
  theme_classic()+
  theme(axis.line = element_line(size = 1))
```
# Arranging and exporting plots

```{r, message=FALSE}
library(gridExtra)
```

```{r}
spp_weight_boxplot <- ggplot(data = surveys_complete, 
                             aes(x = species_id, y = weight)) +
  geom_boxplot() +
  labs(x = "Species", 
       y = expression(log[10](Weight))) +
  scale_y_log10() + 
  labs()

spp_count_plot <- ggplot(data = yearly_counts, 
                         aes(x = year, y = n, color = genus)) +
  geom_line() + 
  labs(x = "Year", y = "Abundance")

grid.arrange(spp_weight_boxplot, spp_count_plot, ncol=2, widths= c(4,6))
```
```{r}
my_plot <- ggplot(data = yearly_sex_counts, 
                  aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus)) +
    labs(title = "Observed genera through time",
        x = "Year of observation",
        y = "Number of individuals") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90,
                                     hjust = 0.5, vjust = 0.5),
          axis.text.y = element_text(colour = "grey20", size = 12),
          text = element_text(size = 16))

ggsave(filename = "Observed genera through time.png", plot = my_plot, width = 15, height = 10)
```


```{r}
combo_plot <- grid.arrange(spp_weight_boxplot, spp_count_plot, ncol = 2, 
                           widths = c(4, 6))

ggsave(filename = "combo_plot_abun_weight.png", plot = combo_plot, width = 15, height = 10, dpi=300)
```

