---
title: "Lab3B"
output:
  html_document: default
  theme:  readable
---
```{r, echo=FALSE, message=FALSE}
library(tidyverse)

```

```{r, message=FALSE}
surveys <- read_csv("data/portal_data_joined.csv")

```

### **Challenge 1**

Using pipes, subset the surveys data to include animals collected before 1995 and
retain only the columns year , sex , and weight .

```{r}
surveys %>% 
  filter(year < 1995) %>% 
  select(year , sex , weight)
```

### **Challenge 2**

Create a new data frame from the surveys data that meets the following criteria:
contains only the species_id column and a new column called hindfoot_cm
containing the hindfoot_length values converted to centimeters. In this
hindfoot_cm column, there are no NA s and all values are less than 3.
Hint: think about how the commands should be ordered to produce this data frame!

```{r}
new_surveys <- surveys %>% 
                mutate("hindfoot_cm"=hindfoot_length/10) %>% 
                filter(!is.na(hindfoot_cm)) %>% 
                filter(hindfoot_cm < 3)%>%
                select(species_id, hindfoot_cm)
new_surveys
```



### **Challenge 3**

1. How many animals were caught in each plot_type surveyed?

```{r}
surveys %>% 
  count(plot_type)
```

2. Use group_by() and summarize() to find the mean, min, and max hindfoot
length for each species (using species_id ). Also add the number of
observations (hint: see ?n ).

```{r, warning=FALSE}
surveys %>% 
  #filter(!is.na(hindfoot_length)) %>% 
  group_by(species_id) %>% 
  summarise(mean_hindfoot=mean(hindfoot_length, na.rm = TRUE),
            min_hindfoot= min(hindfoot_length, na.rm = TRUE),
            max_hindfoot= max(hindfoot_length, na.rm = TRUE),
            number_of_observations= n())
  
```

3. What was the heaviest animal measured in each year? Return the columns
year , genus , species_id , and weight .


```{r}
heaviest_animals_df<-surveys %>% 
  group_by(year) %>% 
  filter(weight==max(weight, na.rm = TRUE)) %>% 
  select(year , genus , species_id , weight)
  
head(heaviest_animals_df,5)
```
# Spreading

Let’s use spread() to transform surveys to find the mean weight of each genus in each
plot over the entire survey period. We use filter() , group_by() and summarise() to
filter our observations and variables of interest, and create a new variable for the
mean_weight .

```{r}
surveys_gw <- surveys %>% 
  filter(!is.na(weight)) %>% 
  group_by(genus,plot_id) %>% 
  summarise(mean_weight=mean(weight))

str(surveys_gw)

```
```{r}
surveys_spread <- surveys_gw %>% 
  spread(key=genus,value = mean_weight)

str(surveys_spread)
```

```{r}
surveys_gw %>% 
  spread(key = genus, value = mean_weight, fill = 0) %>% 
  head()
```
# Gathering
```{r}
surveys_gather <- surveys_spread %>% 
  gather(key = "genus", value = "mean_weight", -plot_id)

str(surveys_gather)
```

```{r}
surveys_spread %>% 
  gather(key="genus", value = "mean_weight", Baiomys:Spermophilus) %>% 
  head()
```
### **Challenge 4**

1. Spread the surveys data frame with year as columns, plot_id as rows, and
the number of genera per plot as the values. You will need to summarize before
reshaping, and use the function n_distinct() to get the number of unique
genera within a particular chunk of data. It’s a powerful function! See
?n_distinct for more.


```{r}
genera_per_plot_df <-surveys %>%
  group_by(plot_id,year) %>% 
  summarise("number_of_genera_per_plot"= n_distinct(genus)) %>% 
  spread(key = year, value = number_of_genera_per_plot)

head(genera_per_plot_df)
```

2. Now take that data frame and gather() it again, so each row is a unique
plot_id by year combination.




```{r}
gathered_df <- genera_per_plot_df%>% 
               gather(key="year", value = "number_of_genera_per_plot",-plot_id)

head(gathered_df)
```

3. The surveys data set has two measurement columns: hindfoot_length and
weight . This makes it difficult to do things like look at the relationship between
mean values of each measurement per year in different plot types. Let’s walk
through a common solution for this type of problem. First, use gather() to
create a dataset where we have a key column called measurement and a
value column that takes on the value of either hindfoot_length or weight .
Hint: You’ll need to specify which columns are being gathered.


```{r}

```

