---
title: "Sequence Alignment, Phylogenetic Analysis and Tree Visualization"
output: html_document
---


```{r, message=FALSE}
library(tidyverse)
library(ggtree)
```



## Visualizing and Annotating PhylogeneticTrees with R+ggtree

Horizontal dimension- genetic change
branches- evolutionary lineages changing over time (longer the branch larger the change)
units- nucleotide substitutions per site

```{r}
tree <- read.tree("data/tree_newick.nwk")
tree

tree %>% ggplot()+
  geom_tree()+
  theme_tree()+
  labs(title="Plotting with ggplot")

tree %>% ggtree()+
  theme_tree2()+
  labs(title="Phylogram")

tree %>% ggtree(branch.length = "none") +labs(title= "Cladogram")

tree %>% ggtree(branch.length = "none", color="blue", size= 2, linetype=3) +
  labs(title= "Cladogram with aes")

```

## ** Exercise 1**

1. Create a slanted phylogenetic tree.
```{r}

tree %>% ggtree(layout = "slanted")+
  theme_tree2()
```
2. Create a circular phylogenetic tree
```{r}
tree %>% ggtree(layout = "circular")+
  theme_tree2()
```

3. Create a circular unscaled cladogram with thick red lines

```{r}
tree %>% ggtree(branch.length = "none", color="red", layout = "circular", linetype=1, size=1)
```

```{r}
p <- tree %>% ggtree()+
  geom_nodepoint()+
  geom_tippoint()+
  geom_tiplab()
print(p)
```
## ** Exercise 2**


```{r}
p <- tree %>% ggtree()+
  geom_nodepoint(alpha=0.5, size=2.5, color="orange")+
  geom_tippoint(shape=23, fill="red", color="red")+
  geom_tiplab(color= "purple")+
  ggtitle("Phylogenetic Aesthetics")
print(p)
```

