---
title: "Population genomics based on highthroughput sequencing (HTS)"
output: html_document
Theme: readable
---

```{r, message=FALSE}
library(vcfR)
library(adegenet)
library(reshape2)
library(ggplot2)
library(tidyverse)
```



```{r}
data(vcfR_example)
vcf
```

## Meta Region

1. Viewing the mta region
```{r}
strwrap(vcf@meta[1:7])

```
2. Summarizing the meta information

  a) Summarizing all information
  
```{r}
queryMETA(vcf)
```

  b) Summarizing information for a particular element
```{r}
queryMETA(vcf, element= "DP")
```

  c) Summarizing information for a particular acronym of a particular element
```{r}
queryMETA(vcf, element = "FORMAT=<ID=DP")
```

## Fix region

Getting information about fix region containing information about each variant

```{r}
head(getFIX(vcf))
```

## The gt region

```{r}
vcf@gt[1:6, 1:4]
```


## Reading the vcfR file

```{r}
vcf <- read.vcfR("data/pinfsc50_filtered.vcf.gz")

head(vcf)
```


## Writing vcf file

```{r}
write.vcf(vcf, "data/myVCFdata_filtered.vcf.gz")
```


## Exercise Part I

1. How would we 􀃝nd more information about
read.vcfR()
?
```{r}
?read.vcfR
```

2. How would we learn what the acronym “AD” stands for?

```{r}
queryMETA(x=vcf, element = "AD")
```

3. We used the head() function to view the first few lines of fix data. How would we view the last few lines of fix data?

```{r}
tail(getFIX(vcf))
```

4) There is a column in the fix portion of the data called QUAL. It is not defined in the meta portion of the data because it is defined in the VCF specification. It stands for ‘quality’. Does QUAL appear useful to us? Why or why not?

```{r}
head(data.frame(vcf@fix),5)

print("'QUAL' column is useful as it says the quality of each variant over all samples")
```

5) How would we query the sample names?

```{r}
colnames(vcf@gt)[colnames(vcf@gt) != "FORMAT"]
```



# Analysis of genomic data

1) Reading the vcfR file

```{r}
#library(vcfR)
vcf <- read.vcfR("data/pinfsc50_filtered.vcf.gz")
vcf
```

2) Converting VCF data to genlight object

```{r}
x <-  vcfR2genind(vcf)
x
```

3) Viewing the genotypes stored

### ***vcfR format*** 
```{r}
gt <- extract.gt(vcf, element = "GT")

gt[c(2,6,18), 1:3]

```

### ***genlight format***

```{r}
t(as.matrix(x))[c(1,5,17), 1:3]
```

4) Setting population data for genlight object
```{r}
pop(x) <- as.factor(c("us", "eu", "us", "af", "eu", "us", "mx", "eu", "eu", "sa", "mx", "sa", "us", "sa",
"Pmir", "us", "eu", "eu"))

popNames(x)
```

5) Setting the ploidy of all samples in genelight object to same ploidy

```{r}
ploidy(x) <- 2
```


6) Generating distance matrices

```{r}
x.dist <- dist(x)
```

Generating distance matrices using poppr package

```{r}
#library(poppr)
#x.dist <- poppr::bitwise.dist(x)
```

7) Locate unusual features in a genome chromR
  
  a)
```{r}

# Find the files

vcf_file <- system.file("extdata", "pinf_sc50.vcf.gz", package = "pinfsc50")

dna_file <- system.file("extdata", "pinf_sc50.fasta", package= "pinfsc50")

gff_file <- system.file("extdata", "pinf_sc50.gff", package = "pinfsc50")



# Input the files

vcf <- read.vcfR(vcf_file, verbose = FALSE)

dna <- ape::read.dna(dna_file, format = "fasta")

gff <- read.table(gff_file, sep="\t", quote ="")


# Create a chromR object

chrom <- create.chromR(name= "Supercontig", vcf = vcf, seq = dna, ann = gff, verbose = TRUE)

chrom
```
  
  b) Plot chromR object
  
```{r}
plot(chrom)
```
  read depth = sum over all samples
  Quality = quality of variants
  
  
  c) Plot a chromo plot- displays the samo information as first plot but it distributes the data along its chromosomal coordinates
```{r}
chromoqc(chrom, dp.alpha = 66)
```
  

  d) Processing chromR objects
  
  Plotting densities using plot function
```{r}
chrom <- proc.chromR(chrom, verbose = TRUE)
```
  
```{r}
plot(chrom)
```
  
  e) Plotting chromoqc plot with variant densities, nucleotide content and where in the reference nucleotides (A,C,GT are)
  
```{r}
chromoqc(chrom, dp.alpha = 66)
```
  
  
```{r}
vcf <- read.vcfR("data/pinfsc50_filtered.vcf.gz", verbose = FALSE)

chrom <- create.chromR(name= "Supercontig", vcf=vcf, seq = dna, ann=gff, verbose = FALSE)

chrom <- proc.chromR(chrom, verbose = FALSE)
chromoqc(chrom, dp.alpha = 66)
```
  read depth in this plot is uniform, lacks large variation.
  mapping quality is relatively constant, variants with low mapping quality is omitted.
  
  
8) Tabular summaries

  a) Summaries based on per variants basis.
  Includes sample size (minus missing data), allele counts, heterozygosity, effective size
```{r}
head(chrom@var.info)
```

  b) Summaries based on per windows.
  Includes nucleotide content, genic sites and the number of variants per window (including missing data)
  
```{r}
head(chrom@win.info)
```
  9) Estimating genetic differentiation
  
  Fst- Population differentiation.
  Heterozygosity = the probability that 2 allelles randomly chosen from a population will be different
  
```{r}
data(vcfR_example)
pop <- as.factor(c("us", "eu", "us", "af", "eu", "us", "mx", "eu", "eu", "sa", "mx", "sa", "us", "sa", "Pm
ir", "us", "eu", "eu"))

myDiff <- genetic_diff(vcf, pops = pop, method="nei")

knitr::kable(head(myDiff[, 1:15]))
```
  
```{r}
knitr::kable(head(myDiff[, 16:19]))
```
  
```{r}
knitr::kable(round(colMeans(myDiff[,c(3:9,16,19)], na.rm=TRUE), digits=3))
```
   summarizing data using violin plots
   
```{r}
dpf <- melt(myDiff[,c(3:8,19)], varnames=c("Index", "Sample"), value.name = "Depth", na.rm = TRUE)

p<- ggplot(dpf, aes(x=variable, y=Depth))+ geom_violin(fill="#2ca25f", adjust= 1.2)

p <- p+
  xlab("")+
  ylab("")+
  theme_bw()

p

```


### ***Exercises Part II***

1) You actually have everything you need to make a Manhattan plot. Can you figure out how to plot G′ST (y-axis) by genomic position (POS)?

```{r}
p <-myDiff %>% 
  mutate(POS=as.numeric(POS)/1000) %>% 
  filter(!is.na("Gprimest")) %>% 
  ggplot(aes(x=POS, y=Gprimest))+
  geom_point(colour="blue", alpha= 0.2)+
  #geom_violin(fill="#2ca25f")+
  labs(x="Genomic position (kb)", y="G prime ST")+
  theme_classic()

p
```

2) This Manhatttan plot shouldlook a bit unusual. Can you think of anything that may be wrong with this analysis?

```{r}
table(pop)
```

3) Can you figure out how to zoom in on a particular region of a chromosome in chromoqc()?

```{r}
chromoqc(chrom=chrom, dp.alpha = 66, xlim = c(2*10^5, 4*10^5))
```

4) Can you use the function queryMETA() to look for other data in your file that may be of interest?

```{r}
queryMETA(vcf, element = "INFO=<ID=MQRankSum")
```

