---
title: "Lab 5"
output: html_document
theme: readable
---

```{r, message=FALSE}
library(tidyverse)
```

```{r}
Confirmed_State_3_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-13-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State") %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```

```{r}
str(Confirmed_State_3_13)
```
```{r}
Confirmed_State_9_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv")) %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```
```{r}
str(Confirmed_State_9_13)
```
```{r}
setdiff(Confirmed_State_9_13$Province_State, Confirmed_State_3_13$Province_State)
```

```{r}
Confirmed_State_9_13 <- Confirmed_State_9_13 %>% 
  filter(Province_State != "Recovered")
```


```{r}
Confirmed_State_3_13_9_13_joined <- full_join(x=Confirmed_State_3_13, y=Confirmed_State_9_13, by=c("Province_State"))

head(Confirmed_State_3_13_9_13_joined )
```
```{r}
tail(Confirmed_State_3_13_9_13_joined, 5)
```
```{r}
which(is.na(Confirmed_State_3_13_9_13_joined))
```
```{r}
Confirmed_State_3_13_9_13_joined <- Confirmed_State_3_13_9_13_joined %>% 
  rename(Confirmed_3_13_2020= "Confirmed.x", Confirmed_9_13_2020="Confirmed.y") %>% 
  select(-Country_Region.x, -Country_Region.y) %>% 
  replace_na(list(Confirmed_3_13_2020 = 0)) 
head(Confirmed_State_3_13_9_13_joined)
```

```{r}
which(is.na(Confirmed_State_3_13_9_13_joined))
```

# **Switching between wide and long table formats**

```{r}
Confirmed_State_3_13_9_13_joined_long <- Confirmed_State_3_13_9_13_joined %>% 
              pivot_longer(-c(Province_State),
                            names_to = "Date", values_to = "Confirmed")
```


```{r, fig.width= 5, fig.height= 10}
ggplot(data=Confirmed_State_3_13_9_13_joined_long, mapping = aes(x=Confirmed, y=Province_State))+ geom_point(aes(color=Date))
```

# **Working with time series data**

```{r}
time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")
```
```{r}
 download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", 
               destfile = "data/time_series_covid19_confirmed_global.csv")
```

```{r}
time_series_confirmed <- read_csv(file = "data/time_series_covid19_confirmed_global.csv") %>% 
  rename(Province_State="Province/State", Country_Region="Country/Region")
```
```{r}
head(time_series_confirmed)
```
As noted above this data is in wide format. To convert to long format
```{r}
time_series_confirmed_long  <- time_series_confirmed %>% 
  pivot_longer(cols= -c(Province_State,Country_Region, Lat, Long), names_to = "Date", values_to="Confirmed" )
```


```{r}
head(time_series_confirmed_long)
```


Create another table with the deaths and then join the two tables. 

```{r}
download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", 
               destfile = "data/time_series_covid19_deaths_global.csv")
```

```{r}
time_series_deaths <- read_csv(file="data/time_series_covid19_deaths_global.csv") %>% 
  rename(Province_State= "Province/State", Country_Region="Country/Region")
```
```{r}
time_series_deaths_long <- time_series_deaths %>% 
  pivot_longer(cols=-c(Province_State,Country_Region, Lat, Long), names_to = "Date", values_to="Deaths")

head(time_series_deaths_long)
```

## **Join time series tables**


```{r}
time_series_confirmed_long <- time_series_confirmed_long %>% unite(col=Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)

head(time_series_confirmed_long)
```

```{r}
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(col=Key, Province_State, Country_Region, Date, sep = ".") %>% select(Key, Deaths)
```


```{r}
time_series_long_joined <- full_join(time_series_confirmed_long, time_series_deaths_long, by= c("Key")) %>% 
  select(-Key)
head(time_series_long_joined)
```

```{r}
which(is.na(time_series_long_joined$Confirmed))
```

```{r}
which(is.na(time_series_long_joined$Deaths))
```

```{r}
library(lubridate)
```

```{r}
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
```

```{r}
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(cols= c(Confirmed, Deaths), names_to="Report_Type", values_to="Counts")

head(time_series_long_joined_counts)
```

# **Making Graphs from the time series data**

```{r}
time_series_long_joined %>% 
  group_by(Country_Region, Date) %>% 
  #summarise("Confirmed" =sum(Confirmed), "Deaths"=sum(Deaths)) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter(Country_Region=="US") %>% 
  ggplot(aes(x=Date, y= Deaths))+
  geom_point()+
  geom_line()+
  ggtitle ("US COVID-19 Deaths")
```

look at the US data in the context of a few other countries

```{r}
time_series_long_joined %>% 
  group_by(Country_Region, Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>%
  filter(Country_Region %in% c("China", "Japan", "Korea, South", "Italy", "Spain", "US")) %>%
  ggplot(mapping=aes(x=Date, y=Deaths))+
  geom_point()+
  geom_line()+
  ggtitle("COVID-19 Deaths")+
  facet_wrap(~Country_Region, ncol=2, scales = "free_y")
  
  
```


```{r}
time_series_long_joined %>% 
  group_by(Country_Region, Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>%
  filter(Country_Region %in% c("China", "Japan", "Korea, South", "Italy", "Spain", "US")) %>%
  ggplot(mapping=aes(x=Date, y=Deaths, color=Country_Region))+
  geom_point()+
  geom_line()+
  ggtitle("COVID-19 Deaths")

```

Use the alternative data frame with column Report_Type to show Confirmed and Deaths



```{r}
time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>%
  summarise(Counts=sum(Counts)) %>% 
  filter(Country_Region=="US") %>% 
  ggplot(aes(x=Date, y= log2(Counts), fill=Report_Type, color=Report_Type))+
  geom_point()+
  geom_line()+
  ggtitle("US COVID-19 Cases")
```

## **Exercise**

1. Uploaded the answers in [Lab 5A](Lab_5A.html)
2. Revise the above example for joining tables using 6/13/2020 and 9/13/2020. Plot the data as a bar plot. Make sure your graph looks nice in the report by adjusting the height of the graph in the R code chunk header.

```{r, fig.width= 10, fig.height= 10}
Confirmed_State_3_13_9_13_joined_long %>% 
  ggplot(mapping = aes(x=Confirmed, y=Province_State))+ geom_bar(aes(fill=Date, group=Date), stat = "identity", position = position_dodge())+
  theme_classic()
```



3. Add more informative x and y labels and a title to the above graph in Ex2.

```{r, fig.width= 10, fig.height= 10}
Confirmed_State_3_13_9_13_joined_long %>% 
  ggplot(mapping = aes(x=Confirmed, y=Province_State))+
  geom_bar(aes(fill=Date, group=Date), stat = "identity", position = position_dodge()) +
  labs(title = "COVID 19 Cases Comparisons",
       x= "Number of Confirmed Cases",
       y= "State/Province")+
  theme_classic()
```







