---
title: "Programming in R - Control Structures and Functions"
output: html_document
Theme: readable
---


# Control Structures

## Conditional statements with if-else

```{r}
x <- runif(1,0,10)
if (x>3){
  y <- 10
} else {
  y<-0
}

print(paste("x =", x))
print(paste("y =", y))
```

## Important Operators

### for Loops

```{r}
for(i in 1:10){
  print(i)
}
```


```{r}
x <- c("a", "b", "c", "d")
for (i in 1:4){
  print(x[i])
}
```


### seq_along

```{r}
for (i in seq_along(x)){
  print(paste("object = ", x[i], ", index = ", i, sep=""))
}
```


```{r}
for (letter in x){
  print(paste("object = ", letter, ", index = ", which(x==letter), sep=""))
}
```

 for one line loop the curly braces are not necessary
```{r}
for (i in 1:4) print(paste("object is ", x[i], "index is", i ,sep=" "))
```

### Nested for loops

Nesting beyond 2 to 3 levels often makes it dicult to read/understand the code
```{r}
x<- matrix(1:6, 2,3)
for (i in seq_len(nrow(x))){
  for(j in seq_len(ncol(x))){
    print(paste ("row =", i, ", column = ", j, ", value = ",x[i,j], sep=""))
  }
}
```


### while Loops

```{r}
count <- 0
while(count<10){
  print(paste("currentcount = ", count, ", next count = ", (count+1), sep=""))
  count <- count+1 
}
```
Preventing while loops resulting in an infinite loop

```{r}
z<- 5
set.seed(1)
while(z >= 3 && z <= 10) {
        coin <- rbinom(1, 1, 0.5)
        print(paste("coin = ", coin, sep=""))
        if(coin == 1) {  ## random walk
                z <- z + 1
        } else {
                z <- z - 1
        } 
}
print(z)
```

## Functions

```{r}
f <- function(){
  print("This is supposed to be an ampty function")
}

f()

class(f)
```


```{r}
 f <- function() {
         cat("Hello, world!\n")
 }
f()
```


```{r}
f <- function(num){
  for (i in seq_len(num)){
    cat("Current number = ", i, "Hello, world!\n")
  }
}

f(5)
```

Last value of the function is the return value
```{r}
f <- function(num){
  hello <- "Hello, world!\n"
  for (i in seq_len(num)){
    cat(i, hello)
  }
  
  chars <- nchar(hello) * num
  chars
}
meaningoflife <- f(5)
print(meaningoflife)
```

Setting a default value to an argument in a function

```{r}
f <- function(num=1){
  hello <- "Hello, world!\n"
  for (i in seq_len(num)){
    cat(i, hello)
  }
  
  chars <- nchar(hello) * num
  chars
}
meaningoflife <- f()
print(meaningoflife)
```

### *** Obtaining the formal argument, the arguments included in the function definition

```{r}
formals(f)
```

```{r}
f(num=2)
```

