---
title: "Programming in R - Regular Expressions"
output: html_document
Theme: readable
---

```{r, message=FALSE}
library(stringr)
library(tidyverse)
```


# Basic Matches

1. Matching exact strings
```{r, message=FALSE}
x<- c("apple", "banana", "pear")

str_view(x,"an")


```
2. Matching any character.
shows only first match in a string). Here, any thing that surrounds the letter a wil be matched 
```{r, message=FALSE}
str_view(x, ".a.")
```
3. Matching a "." (dot) 

```{r, message=FALSE}
dot <- "\\."
writeLines(dot)
x<- c("abc", "a.c", "bef")
str_view(x, "a\\.c")
```
4. Matching a "\" (escape)

```{r, message=FALSE}
x<- "a\\b"
writeLines(x)
str_view(x, "\\\\")
```

### **Exercise 14.3.1.1**

1. Explain why each of these strings don’t match a \: "\", "\\", "\\\".

```{r, message=FALSE}

writeLines("Answer-\\  escapes special behaviour of \\ hence wont match \\. \\\\  escapes special behaviour and escapes symbol. \\\\\\ is three slashes and wont match \\")
```

2. How would you match the sequence "'\?

```{r, message=FALSE}
x<- x<- "\'\\"
writeLines(x)
```

3. What patterns will the regular expression \..\..\.. match? How would you represent it as a string?

```{r, message=FALSE}
x<- "\\..\\..\\.."
writeLines(x)
```

# Anchors

4. Matching from the start of a string

eg: Matching words starting with "a"

```{r}
x<- c("apple", "banana", "pear")

str_view(x,"^a")
```
5. Matching from the end of a string. 

eg: Matching words ending with "a"

```{r}
str_view(x, "a$")
```
6. Matching a complete string from start to end.
eg: matching "apple"

```{r}
x<- c("apple pie", "apple", "apple cake" )
str_view(x, "apple")
str_view(x, "^apple$")
```
### ***Exercise***

1. How would you match the literal string "$^$"?
```{r}
x <- "$^$"
str_view(x, "\\$\\^\\$")
```
2. Given the corpus of common words in stringr::words, create regular expressions that find all words that:
Start with “y”.



```{r}
str_view(stringr::words, "^y", match = TRUE)
```

End with “x”


```{r}
str_view(stringr::words, "x$", match = TRUE)
```


Are exactly three letters long. (Don’t cheat by using str_length()!)

```{r}
str_view(stringr::words, "^...$", match = TRUE)
```


Have seven letters or more.

```{r}
str_view(stringr::words, "^.......", match = TRUE)
```

# Character classes and alternatives

7. Matching a dot containing word using charcater class rather than "\\."

```{r}
x<- c("abc", "a.c", "a*c", "a c")
str_view(x,"a[.]c")
```


```{r}
str_view(x, ".[*]c")
```

```{r}
str_view(x, "a[ ]")
```

8. Picking between one or more alternative patterns

```{r}
x <- c("grey", "gray", "grty")
str_view(x,"gr(e|a|t)y")
```

### ***Exercise 14.3.3.1***

1. Create regular expressions to find all words that:

Start with a vowel.
```{r}
str_view(stringr::words, "^[aeiou]", match = TRUE)
```
That only contain consonants. (Hint: thinking about matching “not”-vowels.)

```{r}
str_view(stringr::words, "^[^aeiou]", match = TRUE)
```

End with ed, but not with eed.

```{r}
str_view(stringr::words, "[^e]ed$", match = TRUE)
```
End with ing or ise

```{r}
str_view(stringr::words, "(ing|ise)$", match=TRUE)
```
2. Empirically verify the rule “i before e except after c”.
```{r}
 str_view(stringr::words, "ie & ^ci", match = TRUE)
```

3. Is “q” always followed by a “u”?

Answer-Yes
```{r}
str_view(stringr::words, "uq", match = TRUE)
```

4. Write a regular expression that matches a word if it’s probably written in British English, not American English

```{r}

```


5. Create a regular expression that will match telephone numbers as commonly written in your country.

```{r}
x <- c(4133333413, 8457777777, 413) 
str_view(x, "^413.......")
```
# Repetition

9. Matching a pattern 0 or 1 times.
eg: Match "CC" in the string x

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII,  CCCCCLLLLXXXX, ccCCCLX"
str_view(x, "CC?")

```

10. Matching a pattern 1 or more times

eg: Match "CC" in x

```{r}
str_view(x, "CC+")
```

11. Matching a pattern 0 or more time.
eg: Matching "CLX" in x

```{r}
str_view(x, "C[LX]+")
```

12. Matching exactly n matches

eg: Match "C" twice in x

```{r}
str_view(x, "C{2}")
```

13. Match n or more matches
eg: Match more than 2C s in x

```{r}
str_view(x, "C{2,}")
```


```{r}
str_view(x,"C{2,}?")
```
Matching a string between two repititions 
eg: matching "C" between 2-3 repititions

```{r}
str_view(x, "C{2,3}")
```


```{r}
str_view(x, "C[LX]+?")
```


### ***Exercises***

1. Describe the equivalents of ?, +, * in {m,n} form

? = {0,1}
+ = {1, }
* = {0,}


2. Describe in words what these regular expressions match: (read carefully to see if I’m using a regular expression or a string that defines a regular expression.)
^.*$- word exactly starting with a dot and ends with star
"\\{.+\\}"
\d{4}-\d{2}-\d{2}
"\\\\{4}"

```{r}
writeLines("^.*$")
writeLines("\\{.+\\}")

writeLines("\\\\{4}")
```

3.Create regular expressions to find all words that:

Start with three consonants.

```{r}
str_view(stringr::words, "^[^aeiou]{3}", match = TRUE)
```

Have three or more vowels in a row.

```{r}
str_view(stringr::words, "[aeiou]{3,}", match = TRUE)
```
Have two or more vowel-consonant pairs in a row.

```{r}
#str_view(, "[aeiou][^aeiou]+?", match = TRUE)
str_view(stringr::words, "([aeiou][^aeiou]){2,}", match = TRUE)
```

## Grouping and References

14. Capture a string  and refer to the same string as with back references like \1,\2

eg: capture any four letters where third and fourth letters are similar to first and second letters

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

### ***14.3.5.1 Exercises#

Describe, in words, what these expressions will match:

(.)\1\1-Detects same letter 3 times in a row of a string

```{r}
str_view("bbbaaajjjjj", "(.)\\1\\1", match = TRUE)
```
"(.)(.)\\2\\1" = matches 4 letters where second two letters are the inverse of the first two


```{r}
str_view(stringr::words, "(.)(.)\\2\\1", match = TRUE)
```
(..)\1- match four letters where first two letter are equal to last two letters

```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

"(.).\\1.\\1"- select five letters where 3rd and 5th equal to first letter

```{r}
str_view(stringr::words, "(.).\\1.\\1", match=TRUE)
```

"(.)(.)(.).*\\3\\2\\1"- match 7 plus letters where last three is inverse to first three letters

```{r}
str_view(stringr::words, "(.)(.)(.).*\\3\\2\\1", match=TRUE)
```
2. Construct regular expressions to match words that:

Start and end with the same character.

```{r}
str_view(stringr::words, "^(.).*\\1$", match = TRUE)
```


Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)

```{r}
str_view(stringr::words, "(..).*\\1$", match = TRUE)
```


Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)

```{r}
str_view(stringr::words, "(.).*\\1.*\\1", match = TRUE)
```

## Tools
### Detect matches

15. Detecting a patteren in a string

```{r}
x <- c("apple", "banana", "pear")

str_detect(string = x, pattern = "e")
```

How many common words start with t?

```{r}
sum(str_detect(words,"^t"))
```

What proportion of common words end with a vowel?

```{r}
mean(str_detect(words, "[aeiou]$"))
```


Find all words containing at least one vowel, and negate

```{r}
no_vowels_1 <- !str_detect(words, "[aeiou]")
```


Find all words consisting only of consonants (non-vowels)

```{r}
no_vowels_2 <- str_detect(words, "^[^aeiou]+$")
```


```{r}
identical(no_vowels_1, no_vowels_2)
```

16. selecting elements that match a pattern

```{r}
words[str_detect(words, "x$")]
```


```{r}
str_subset(words, "x$")
```

17. Extracting strings from a column of a data frame/tibble

```{r}

df <- tibble(
  word =words,
  i = seq_along(word)
)


df %>% filter(str_detect(word, "x$"))
```
18. Counting the number of matches

```{r}
x <- c("apple", "banana", "pear")

str_count(x, "a")

```

On average how many vowels per word

```{r}
mean(str_count(words, "[aeiou]"))

```



```{r}
df %>% 
  mutate(vowels=str_count(word, "[aeiou]"),
         consonants= str_count(word, "[^aeiou]"))
```

```{r}
str_count("abababa", "aba")
```

```{r}
str_view_all("abababa", "aba")
```


### 14.4.1.1 Exercises

For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls.

Find all words that start or end with x.

```{r}

reg <- words[str_detect(words, "^x|x$")]


starting_x <-str_detect(words, "^x")

ending_x <- str_detect(words, "x$")

sep_words <- words[starting_x|ending_x]

identical(reg, sep_words)
```


Find all words that start with a vowel and end with a consonant.


```{r}
reg <-words[str_detect(words, "^[aeiou]|[^aeiou]$")]


starting_vowel<- str_detect(words, "^[aeiou]")

ending_cons <- str_detect(words, "[^aeiou]$")

sep_words <- words[starting_vowel|ending_cons]

identical(reg, sep_words)

```


Are there any words that contain at least one of each different vowel?

```{r}

sep_words<- str_detect(words, "[a]+[e]+[i]+[o]+[u]+")

sep_words <-  str_detect(words, "[a]{1,}") &
        str_detect(words, "[e]{1,}") &
        str_detect(words, "[i]{1,}") &
        str_detect(words, "[o]{1,}") &
        str_detect(words, "[u]{1,}")





identical(reg, sep_words)
```



What word has the highest number of vowels? 

```{r}
sum_lens<- str_count(words, "[aeiou]")

words[which(sum_lens == max(sum_lens))]


```


What word has the highest proportion of vowels? (Hint: what is the denominator?)

```{r}
tot_len <- str_count(words, "")
prop <- sum_lens/tot_len

words[which(prop == max(prop))]
```

