---
title: "Microbiome Analysis usingphyloseq"
output: html_document
---


# **Ordination**

```{r, message=FALSE}
library(phyloseq)
library(ggplot2)
library(plyr)
```

```{r}
data(GlobalPatterns)
theme_set(theme_bw())
```


Removing OTUs that do not show appear more than 5 time in more than half the samples
```{r}
GP = GlobalPatterns
wh0 = genefilter_sample(GP, filterfun_sample(function(x) x>5), A=0.5*nsamples (GP))

GP1 = prune_taxa(wh0, GP)
                        
```

Transforming to even sampling depths

```{r}
GP1 = transform_sample_counts(GP1, function(x) 1E6 * x/sum(x))
```

Keep only the most abundant five phyla

```{r}
phylum.sum = tapply(taxa_sums(GP1), tax_table(GP1)[, "Phylum"], sum, na.rm=TRUE)
top5phyla = names(sort(phylum.sum, TRUE))[1:5]
GP1 = prune_taxa((tax_table(GP1)[, "Phylum"] %in% top5phyla), GP1)
```

Define a human-associated versus non-human categorical variable

```{r}
human = get_variable(GP1, "SampleType") %in% c("Feces", "Mock", "Skin", "Tongue")
sample_data(GP1)$human <- factor(human)
```


### Four main ordination plots

**1) Plotting only OTUs**
```{r}
GP.ord <- ordinate(GP1, "NMDS", "bray")
p1 = plot_ordination(GP1, GP.ord, type="taxa", color="Phylum", title="taxa")
print(p1)
```



```{r}
p1 + facet_wrap(~Phylum, 3)
```

**2) Plotting only samples**

```{r}
p2 = plot_ordination(GP1, GP.ord, type="samples", color="SampleType", shape="human") 
p2 + geom_polygon(aes(fill=SampleType)) + geom_point(size=5) + ggtitle("samples")
```

**3)  Plotting both the samples and OTUs in one “biplot”**

```{r}
p3 = plot_ordination(GP1, GP.ord, type="biplot", color="SampleType", shape="Phylum", title="biplot")
# Some stuff to modify the automatic shape scale
GP1.shape.names = get_taxa_unique(GP1, "Phylum")
GP1.shape <- 15:(15 + length(GP1.shape.names) - 1)
names(GP1.shape) <- GP1.shape.names
GP1.shape["samples"] <- 16
p3 + scale_shape_manual(values=GP1.shape)
```

**4) Plotting samples/OTUs side by side**

```{r}
p4 = plot_ordination(GP1, GP.ord, type="split", color="Phylum", shape="human", label="SampleType", title="split") 
p4
```


```{r}
gg_color_hue <- function(n){
    hues = seq(15, 375, length=n+1)
    hcl(h=hues, l=65, c=100)[1:n]
}
color.names <- levels(p4$data$Phylum)
p4cols <- gg_color_hue(length(color.names))
names(p4cols) <- color.names
p4cols["samples"] <- "black"
p4 + scale_color_manual(values=p4cols)
```


### Supported Ordination Methods

```{r}
dist = "bray"
ord_meths = c("DCA", "CCA", "RDA", "DPCoA", "NMDS", "MDS", "PCoA")
plist = llply(as.list(ord_meths), function(i, physeq, dist){
        ordi = ordinate(physeq, method=i, distance=dist)
        plot_ordination(physeq, ordi, "samples", color="SampleType")
}, GP1, dist)
```
```{r}
names(plist) <- ord_meths

pdataframe = ldply(plist, function(x){
    df = x$data[, 1:2]
    colnames(df) = c("Axis_1", "Axis_2")
    return(cbind(df, x$data))
})
names(pdataframe)[1] = "method"


p = ggplot(pdataframe, aes(Axis_1, Axis_2, color=SampleType, shape=human, fill=SampleType))
p = p + geom_point(size=4) + geom_polygon()
p = p + facet_wrap(~method, scales="free")
p = p + scale_fill_brewer(type="qual", palette="Set1")
p = p + scale_colour_brewer(type="qual", palette="Set1")
p
```
```{r}
plist[[2]]
```


```{r}
p = plist[[2]] + scale_colour_brewer(type="qual", palette="Set1")
p = p + scale_fill_brewer(type="qual", palette="Set1")
p = p + geom_point(size=5) + geom_polygon(aes(fill=SampleType))
p
```

### MDS (“PCoA”) on Unifrac Distances

```{r}
ordu = ordinate(GP1, "PCoA", "unifrac", weighted=TRUE)
plot_ordination(GP1, ordu, color="SampleType", shape="human")
```


```{r}
p = plot_ordination(GP1, ordu, color="SampleType", shape="human")
p = p + geom_point(size=7, alpha=0.75)
p = p + scale_colour_brewer(type="qual", palette="Set1")
p + ggtitle("MDS/PCoA on weighted-UniFrac distance, GlobalPatterns")

```



# **Alpha diversity graphics**

```{r}
library(phyloseq)
library(ggplot2)
```


```{r}
data("GlobalPatterns")

theme_set(theme_bw())
pal = "Set1"
scale_colour_discrete <-  function(palname=pal, ...){
  scale_colour_brewer(palette=palname, ...)
}
scale_fill_discrete <-  function(palname=pal, ...){
  scale_fill_brewer(palette=palname, ...)
}


GP <- prune_species(speciesSums(GlobalPatterns) > 0, GlobalPatterns)
```



```{r}
plot_richness(GP)
```

Plotting alpha diversity measure of interest
```{r}
plot_richness(GP, measures=c("Chao1", "Shannon"))
```
Organize samples along x axis

```{r}
plot_richness(GP, x="SampleType", measures=c("Chao1", "Shannon"))
```


```{r}
sampleData(GP)$human <- getVariable(GP, "SampleType") %in% c("Feces", "Mock", "Skin", "Tongue")
```

```{r}
plot_richness(GP, x="human", color="SampleType", measures=c("Chao1", "Shannon"))
```

```{r}
GPst = merge_samples(GP, "SampleType")
# repair variables that were damaged during merge (coerced to numeric)
sample_data(GPst)$SampleType <- factor(sample_names(GPst))
sample_data(GPst)$human <- as.logical(sample_data(GPst)$human)

p = plot_richness(GPst, x="human", color="SampleType", measures=c("Chao1", "Shannon"))
p + geom_point(size=5, alpha=0.7)
```

```{r}
p$layers

p$layers <- p$layers[-1]
p + geom_point(size=5, alpha=0.7)
```


# **Heatmap Plots**

```{r}
library(phyloseq)
library(ggplot2)
theme_set(theme_bw())
```


 Plotting a 300-taxa data set
 
```{r}
data("GlobalPatterns")
gpt <- subset_taxa(GlobalPatterns, Kingdom=="Bacteria")
gpt <- prune_taxa(names(sort(taxa_sums(gpt),TRUE)[1:300]), gpt)
plot_heatmap(gpt, sample.label="SampleType")

```
 
 Subset a smaller dataset based on an Archaeal phylum
 

```{r}
gpac <- subset_taxa(GlobalPatterns, Phylum=="Crenarchaeota")
```


Default heatmap settings

```{r}
plot_heatmap(gpac)
```

Re-label by a sample variable and taxonomic family

```{r}
(p <- plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family"))
```

Re-lable axis titles

```{r}
p$scales$scales[[1]]$name <- "My X-Axis"
p$scales$scales[[2]]$name <- "My Y-Axis"
print(p)
```

Changing the color scheme

```{r}
plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family", low="#000033", high="#CCFF66")
```

1) dark-blue to red scheme

```{r}
plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family", low="#000033", high="#FF3300")
```

2) Very dark blue to very light blue scheme 

```{r}
plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family", low="#000033", high="#66CCFF")
```
3) Dark on light color scheme

```{r}
plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family", low="#66CCFF", high="#000033", na.value="white")
```

```{r}
plot_heatmap(gpac, "NMDS", "bray", "SampleType", "Family", low="#FFFFCC", high="#000033", na.value="white")
```

### Different ordination methods and distances

1) Plotting NMDS ordination on the jaccard distance
```{r}
plot_heatmap(gpac, "NMDS", "jaccard")
```

2) Plotting detrended correspondence analysis

```{r}
plot_heatmap(gpac, "DCA", "none", "SampleType", "Family")
```

3) Plotting unconstrained redundancy analysis (Principle Components Analysis, PCA)

```{r}
plot_heatmap(gpac, "RDA", "none", "SampleType", "Family")
```

4) Plotting PCoA/MDS ordination on the (default) bray-curtis distance

```{r}
plot_heatmap(gpac, "PCoA", "bray", "SampleType", "Family")
```

5) Plotting MDS/PCoA ordination on the Unweighted-UniFrac distance.

```{r}
plot_heatmap(gpac, "PCoA", "unifrac", "SampleType", "Family")
```

6) Plotting weighted-UniFrac distance and MDS/PCoA ordination

```{r}
plot_heatmap(gpac, "MDS", "unifrac", "SampleType", "Family", weighted=TRUE)
```
7) Plotting heatmaps with R base graphics

```{r}
heatmap(otu_table(gpac))
```
# **Plot Microbiome Network**


```{r}
library(phyloseq)
library(ggplot2)
```


```{r}
data("enterotype")
set.seed(711L)
enterotype= subset_samples(enterotype, !is.na((enterotype)))
```

The plot_net function

```{r}
plot_net(enterotype, maxdist = 0.4, point_label = "Sample_ID")
```

```{r}
plot_net(enterotype, maxdist = 0.3, color = "SeqTech", shape="Enterotype")
```

The plot_network function

```{r}
ig <- make_network(enterotype, max.dist=0.3)

plot_network(ig, enterotype)


```


```{r}
plot_network(ig, enterotype, color="SeqTech", shape="Enterotype", line_weight=0.4, label=NULL)
```
Lowering the  maximum distance decreasing the number of edges in the network

```{r}
ig <- make_network(enterotype, max.dist=0.2)
plot_network(ig, enterotype, color="SeqTech", shape="Enterotype", line_weight=0.4, label=NULL)
```

Replace the Jaccard (default) distance method with Bray-Curtis
```{r}
ig <- make_network(enterotype, dist.fun="bray", max.dist=0.3)
plot_network(ig, enterotype, color="SeqTech", shape="Enterotype", line_weight=0.4, label=NULL)
```

