---
title: "Lab 5A"
output: html_document
theme: readable
---

```{r, message=FALSE}
library(tidyverse)

```



# Boxplot

```{r}
surveys_complete <- read_csv("Results/surveys_complete.csv")

ggplot(data=surveys_complete, mapping=aes(x=species_id, y= weight))+
  geom_boxplot()
```


```{r}
ggplot(data=surveys_complete, mapping=aes(x=species_id, y= weight))+
  geom_boxplot(alpha=0)+
  geom_jitter(alpha=0.3, color= "tomato")
```

What do you need to change inthe code to put the boxplot in front of the points such that it’s not hidden?



```{r}
ggplot(data=surveys_complete, mapping=aes(x=species_id, y= weight))+
  geom_jitter(alpha=0.3, color= "tomato")+
  geom_boxplot(alpha=0)
```

## **Challenges**

Replace the box plot with a violin plot; see geom_violin()

```{r}
ggplot(data=surveys_complete, mapping=aes(x=species_id, y= weight))+
  geom_jitter(alpha=0.3, color= "tomato")+
  geom_violin()
```



Represent weight on the logscale; see scale_y_log10()


```{r}
ggplot(data=surveys_complete, mapping=aes(x=species_id, y= weight))+
  geom_jitter(alpha=0.3, color= "tomato")+
  geom_violin()+
  scale_y_log10()
```

Create boxplot for hindfoot_length. Overlay the boxplot layer on a jitter layer to show actual measurements

```{r}
ggplot(data=surveys_complete, mapping = aes(x= species_id, y=hindfoot_length))+
  geom_jitter( aes(color=as.factor(plot_id)), alpha=0.3)+
  geom_boxplot(alpha=0)
```
Hint: Check the class for
plot_id
. Consider changing the class of
plot_id
from integer to factor. Why does this change how R makes the graph?

```{r, echo=FALSE}
print("When plot_id is in the class numeric R assign colors to plot_id as a gradient. But When it is a factor each factor is colored as a distinct category")
```

# Ploting time series data

Calculate number of counts per year for each genus

```{r}
yearly_counts <- surveys_complete %>% 
 count(year,genus)
```

```{r}
ggplot(data=yearly_counts, aes(x=year, y=n))+
  geom_line()
```
```{r}
ggplot(data=yearly_counts, aes(x=year, y=n))+
  geom_line(aes(group=genus))
```
```{r}
ggplot(data=yearly_counts, aes(x=year, y=n))+
  geom_line(aes(group=genus, color=genus))
```

# Integrating the pipe operator with ggplot2

```{r}
yearly_counts %>% 
ggplot(aes(x=year, y=n))+
  geom_line(aes(group=genus, color=genus))
```
```{r}
yearly_counts_graph <- surveys_complete %>% 
 count(year,genus) %>%
 ggplot(aes(x=year, y=n))+
  geom_line(aes(group=genus, color=genus))

yearly_counts_graph
```

# Faceting

```{r}
ggplot(data=yearly_counts, aes(x=year, y=n))+
  geom_line(aes(group=genus, color=genus))+
  facet_wrap(facets = vars(genus))
```

```{r}
yearly_sex_counts <- surveys_complete %>% 
  count(year, genus, sex)
```

```{r}
ggplot(data=yearly_sex_counts, mapping = aes(x=year, y= n))+
  geom_line(aes(color= sex))+
  facet_wrap(facets = vars(genus))
```
```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(sex), cols =  vars(genus))
```

```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(rows = vars(genus))
```

```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(cols =  vars(genus))
```


# ggplot2 themes

```{r}
ggplot(data = yearly_sex_counts, 
       mapping = aes(x = year, y = n, color = sex)) +
  geom_line() +
  facet_grid(cols =  vars(genus))+
  theme_bw()
```

## **Challenge**


Use what you just learned to create a plot that depicts how the average weight of each species changes through the years.

```{r, fig.width=10}
surveys_complete %>% 
  group_by(species, year) %>% 
  summarise(Mean_weight= mean(weight)) %>% 
  ggplot(aes(x=year, y=Mean_weight))+
  geom_line(aes(color=species))+
  facet_wrap(facets = vars(species), scales = "free_y")+
  theme_classic()
  
```

```{r, fig.width=10}
ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus))+
  labs(title = "Observed genera through time",
       x="Year of observation",
       y= "Number of individuals")+
  theme_classic()+
  theme(text = element_text(size=16))
```

```{r, fig.width=10}
ggplot(data = yearly_sex_counts, aes(x = year, y = n, color = sex)) +
    geom_line() +
    facet_wrap(vars(genus))+
  labs(title = "Observed genera through time",
       x="Year of observation",
       y= "Number of individuals")+
  theme_classic()+
  theme(axis.title.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
        text = element_text(size=16))
```

